Moral Foundations of U.S. Political News Organizations

### 2019-02-04 pilot_manuscript.tex
updated the text a bit.

  \title{Moral Foundations of U.S. Political News Organizations}
    \author{William E. Padfield\textsuperscript{1}~\& Erin M. Buchanan, Ph.D.\textsuperscript{1}}
    \date{}

\shorttitle{MORAL NEWS}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Missouri State University}
\keywords{politics, morality, psycholinguistics}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

Morality and ethics have been of interest to thinkers, academics, and philosophers since antiquity. Starting chiefly in the twentieth century, a scientific approach to humans' understanding of morality emerged under the domain of psychology. Theories attempting to explain the development and application of people's moral intuitions built the foundation for the subfield of moral psychology. As the field developed, however, considerable debate has taken place regarding operational definitions of \enquote{morality.} This remains an issue in the field in the twenty-first century as researchers attempt to infer moral and political leanings from text and speech.

@@

As a discipline, modern moral psychology started in the late 1960s with Lawrence Kohlberg (Haidt \& Graham, 2007). Kohlberg's research popularized his theory of the development of moral reasoning. This theory establishes the steps of moral reasoning through which humans proceed as they mature. Kohlberg borrowed from Jean Piaget's stages of cognitive development in which children progress from the sensorimotor through to the formal operations stage (Haidt \& Graham, 2007). Similarly, Kohlberg found people typically start with a \enquote{pre-conventional} understanding of morality during infancy and progress through several steps until they reach a \enquote{post-conventional} ethics (Haidt \& Graham, 2007). People who have reached the post-conventional stage are said to be able to weigh competing abstractions and reason their way to a conclusion that promotes justice. From Kohlberg's perspective, issues of justice and fairness comprise the foundation of morality (Haidt \& Graham, 2007). This view persisted until it encountered criticism in the early 1980s.

Kohlberg's conception of morality faced major scrutiny from psychologist Carol Gilligan. In 1982, Gilligan criticized Kohlberg's theory on the grounds that it focused solely on the moral concerns of men, and that it ignored those of women (Haidt \& Graham, 2007). She offered a historic argument contending women have traditionally filled roles related to caring and nurturing. Gilligan thus asserted morality was built upon an alternative moral foundation: caring (Haidt \& Graham, 2007). This debate between competing conceptions of morality did not resolve until Gilligan and Kohlberg conceded the existence of two moral foundations: justice and caring (Haidt \& Graham, 2007). While this new direction in moral psychology appeared to represent a more inclusive outlook on the construct, these novel ideas would soon be challenged on the grounds of its apparent western-centric outlook.

In their brief overview of the history of moral psychology, Graham, Haidt, and Nosek (2009) explained Shweder, Much, Mahapatra, and Park's objections to moral psychology as it stood in the late 1980s. Their criticism centered on the fact moral psychology concerned itself with issues regarding justice and individuals' rights. Such a system, they argued, did not account for moral concerns outside of the western world (Graham et al., 2009). Individually focused concerns can be grouped under an overarching \enquote{ethic of autonomy,} which was thought to be one of three ethics upon which humans base moral decisions. The other two ethics were the \enquote{ethic of community} (comprising one's duty to their family, tribe, etc.) and the \enquote{ethic of divinity}" - representing one's duty not to defile their God-given body and soul (Graham et al., 2009). In the 2000s, Haidt and Graham (2007) took this line of reasoning further in their assertion that moral psychology favored certain political ideologies over others.

@@

The researchers settled on these specific foundations after the completion of a literature survey of research in anthropology and evolutionary psychology (Graham et al., 2011). The researchers attempted to locate virtues and morals corresponding to \enquote{evolutionary thinking.} For instance, the researchers cited Mauss' work on reciprocal gift-giving, which informed the establishment of the \emph{fairness/reciprocity} foundation. Additionally, evolutionary literature on disgust and its correlation to human behavior regarding food and sex informed the \emph{purity/sanctity} foundation (Graham et al., 2011). The researchers identified the five \enquote{top candidates} for the foundations of human cultures' morality (Graham et al., 2011).

@@

In order to capture language's role in moral and political reasoning, Graham et al. (2009) formulated the Moral Foundations Dictionary (MFD) in order to capture moral reasoning and justification as used in speech and text. The MFD is composed of 259 words, with around 50 words assigned to each of the five foundations. The researchers created a preliminary list of words that they believed would be associated with the five foundations. Then, using the Linguistic Inquiry and Word Count (LIWC; Pennebaker, Booth, \& Frances, 2007) computer program, they analyzed transcripts of liberal and conservative Christian sermons in order to obtain frequencies of the occurrence of words from the researchers' initial list. The researchers manually checked the results from LIWC in order to make sure the results make sense given the contexts and rhetorical devices used in the sermons. The researchers offered the following example from a Unitarian sermon as a demonstration of ambiguous statements requiring human verification: \enquote{Don't let some self-interested ecclesiastical or government authority tell you what to believe, but read the Bible with your own eyes and open your heart directly to Jesus} (Graham et al., 2009). This sentence added to the \emph{authority/respect} total in LIWC's analysis, but it appears to suggest that one should reject authority in this context. The researchers eliminated this sentence from the \emph{authority/respect} raw count on account of this discrepancy between the use of authority-related words and the speaker's clear intentions (Graham et al., 2009).

@@

When considering the issue of positive versus negative valence in MFD words, the question of how texts are analyzed vis-a-vis the MFD remains. How can raw percentage of MFD word occurrence capture the valence and focus of the writer or speaker? If 2\% of a politician's speech features positive words (i.e. \enquote{benefit} and \enquote{defend}) from the MFD \emph{harm/care} list, how can researchers be sure the level and nature of the speaker's \enquote{endorsement} of the foundation equals that of another politician whose speech contained negatively connoted MFD words from the \emph{harm/care} list? They would have equal endorsements as far as the numbers are concerned, but the words used and focus given are on opposite sides of the \emph{harm/care} spectrum.

@@

Participants saw the stimulus word and responded on each scale. The valence scale featured a smiling figure at one end (representing pleasantness) and a frowning figure at the other end (for unpleasantness). The arousal scale had a \enquote{wide-eyed} figure at one end with a sleepy figure at the other, representing stimulating and unstimulating respectively. Finally, the control scale featured a large figure, indicating the highest degree of control, at one end and a small figure, indicating a lack of control, at the other end (Bradley \& Lang, 1999). The end result of this procedure yielded affective norms along the three dimensions for 1,040 English words (Bradley \& Lang, 1999). ANEW represented an important first step in establishing affective norms for large numbers of English words. However, later researchers found the 1,040-word list to be limiting for a language consisting of thousands of words.

Warriner, Kuperman, and Brysbaert (2013) exponentially lengthened the list of words with affective norms to 13,915 English lemmas, the base forms of words without inflection (i.e. \enquote{watch} rather than \enquote{watched} and \enquote{watching}). The researchers recognized the importance of affective norms in several areas of study, including emotion, language processing, and memory (Warriner et al., 2013). They argue the list of words included in ANEW is sufficient for small-scale factorial research designs, but the list is \enquote{prohibitively small} for larger-scale \enquote{megastudies} that are common in psycholinguistic research today (Warriner et al., 2013).

In order to source a large number of lemmas for affective ratings, the researchers drew from several validated sources. These include the 30,000 lemmas with age-of-acquisition (average age at which a particular word is learned) ratings gathered by Kuperman, Stadthagen-Gonzalez, and Brysbaert (2012) as well as the content lemmas from the SUBTLEX-US corpus consisting of subtitles from various forms of visual media (Warriner et al., 2013). This data collection resulted in the final list of 13,915 lemmas. Lists of 346-350 words were presented to participants recruited through the Amazon Mechanical Turk subject pool. Participants rated the words along one of the three dimensions, unlike the ANEW project in which participants rated each word along all three dimensions at once. The researchers used a nine-point scale similar to the one used by Bradley and Lang (1999) when collecting ratings for ANEW (Warriner et al., 2013).

The researchers noted several points of interest upon observing ratings. First, they found that valence and dominance ratings had a negative skew, indicating more words elicited feelings of happiness and control than their respective opposites. Also, when examining the relationship between valence and arousal ratings, the researchers found a U-shaped relationship. This indicates words with high degrees of positivity and negativity elicited higher arousal (Warriner et al., 2013). These observations along with the now-greatly expanded list of affective norms has been applied to several lines of inquiry in psycholinguistics.

Warriner and Kuperman (2015) utilized the new affective norms list in order to investigate the validity of the Pollyanna hypothesis, or the prevalence of a generally optimistic outlook in humans as reflected in language. The researchers were able to conclude the existence of a greater number of positive-valence English words in the list of 13,915 lemmas. Additionally, after observing token frequency in a number of text corpora, including SUBTLEX-US, COCA, BNC, TASA, and HAL, the researchers found that words with postie valence were also used more frequently (Warriner \& Kuperman, 2015). While the researchers concede the possibility of an acquiescence bias in ratings as a possible explanation for the observed positivity bias, this investigation represents one application of the Warriner et al. (2013) list in emotional studies.

In addition to applications in emotion research, the Warriner norms (2013) have been utilized in cognitive research as well. One cognition-based study investigates the relationship between emotion and response latencies in word recognition. Kuperman, Estes, Brysbaert, and Warriner (2014) sought to use the Warriner et al. (2013) norms to fill in the knowledge gaps regarding variance in word recognition. The researchers drew several conclusions regarding emotion and word recognition (specifically in naming and lexical decision tasks). First, Kuperman et al. (2014) found slower decision-making and reading times in negative-valence words, faster times in neutral words, and even faster times in words with positive valence. The researchers also concluded that words causing higher arousal tend to have slower decision times than less-arousing words. They found valence had a stronger effect on recognition than arousal (both effects were independent, not interactive). They found an interaction between emotion and word frequency such that valence and arousal are more effective on lower frequency words than high frequency words. Finally, Kuperman et al. (2014) found a greater effect of valence and arousal on response latency for lexical decision tasks than for naming tasks (Kuperman et al., 2014). This research serves as further evidence that the Warriner and Kuperman (2015) list can be used for research inquiries both within and without the field of psycholinguistics.

In the present studies, the researchers used the Warriner et al. (2013) list in order to denote the valence of the words appearing in the news articles scraped from the internet. Valence was considered as another independent variable and its relationship with the words comprising the Moral Foundations Dictionary were of chief interest to the researchers. The valence was used as a means to determine whether individual words in the MFD represented more positive aspects of their respective foundation or if they denoted a more negative aspect of the foundation. Incorporating word valence into a study involving the MFD is meant to alleviate some of the issues regarding the aforementioned ambiguity regarding the words in the Moral Foundations Dictionary.

@@

The use of language both to express and further an ideological goal has been documented in the techniques employed by candidates for political office in the U.S. Druckman, Jacobs, and Ostermeier ((2004)) considered political \enquote{issues} as communication that attempts to persuade constituents to vote for the candidates based on their strengths in matters of public policy. According to the researchers, \enquote{image} priming describes techniques deployed in order to sway votes based on favorable aspects of the candidate's behavior and personality (Druckman et al., 2004).The researchers investigated political issue and image priming on the part of candidates as implied by the disproportionate attention candidates paid to particular issues over others. The researchers found numerous examples of issue and image priming during the 1972 re-election campaign of Richard Nixon. They linked the Nixon administration's awareness of the issues for which the president had public support to the issues he should emphasize (and prime) during the campaign. Likewise the researchers found evidence that Nixon's team was aware of negative evaluations of his warmth and trustworthiness, and thus took steps to prime his purportedly positive qualities, including strength and competence (Druckman et al., 2004).

Druckman et al. (2004) also cited research from Iyengar and Kinder (1987) suggesting the news media affected perceptions of President Jimmy Carter's competence by emphasizing (e.g.~priming) issues related to energy, defense, and the economy. This focus implies news media may affect perception of politicians based on where the media places emphasis. Other research into news media suggests certain media outlets, at least indirectly, may have an effect on the voting records of representatives in Congress (Clinton \& Enamorado, 2014). Specifically, the researchers identified a pattern of declining support for President Bill Clinton's policies chiefly among Republicans in the House of Representatives after the Fox News Channel began broadcasting on cable and satellite systems in their respective districts. As Fox News was, at the time of its launch in 1996, the only outwardly ideological national news network, the researchers were able to track its spread across the country and observe voting records of members of Congress both before and after Fox News' arrival (Clinton \& Enamorado, 2014). The researchers concluded that members of Congress, excluding those newly elected at the time of Fox News Channel's emergence, attempted to anticipate resultant conservative-leaning shifts among their constituents by bolstering their conservative voting record before the next election (Clinton \& Enamorado, 2014).

@@

For Experiment 1, the researchers approached the methodology with the intention to answer a method question. That is, this portion of the current research was conducted in order to solidify the best method by which to analyze political news text under the Moral Foundations Theory framework. The researchers hypothesized the news sources genrally perceived as liberal leaning (\emph{NPR} and \emph{The New York Times}) would contain MFD words and valences indicating endorsements of the individualizing moral foundations (\emph{harm/care} and \emph{fairness/reciprocity}). Additionally, the researchers hypothesized the two sources generally perceived to be conservative leaning (\emph{Fox News} and \emph{Breitbart}) would feature MFD words and valences indicating equal endorsement of all five foundations.

@@

Political articles were scraped from the websites of four notable U.S. news sources. The sources were \emph{The New York Times}, \emph{National Public Radio (NPR)}, \emph{Fox News}, and \emph{Breitbart}. They were selected for their widespread recognition and the fact they are easily categorized (by the general public) according to perceived political lean. In general, \emph{The New York Times} and \emph{NPR} are perceived by many to have a liberal bias or lean. In contrast, Fox News and Breitbart are believed to have a conservative bias or lean (Mitchell, Matsa, Gottfried, \& Kiley, 2014). Political articles in particular were identified and subsequently scraped by including the specific URL directing to each source's political content in the \emph{R} script. For example, rather than scrape from nytimes.com, which would return undesired results (non-political features, reviews, etc.), we instead included nytimes.com/section/politics so that more or less exclusively political content was obtained. All code for this manuscript can be found at OSF LINK, and the scripts are provided inline with this manuscript written with the \emph{papaja} library (Aust \& Barth, 2017).

@@

The article text was processed using the \emph{tm} and \emph{ngram} packages in \emph{R} in order to render the text in lowercase, remove punctuation, and fix spacing issues ({\textbf{???}}). The individual words were then reduced to their stems (i.e., \emph{abused} was stemmed to \emph{abus}). The same procedure was applied to the MFD words and the words in the Warriner et al. (2013) dataset. Using the Warriner et al. (2013) dictionary, the words making up each of the five foundations in the MFD were assigned their respective valence value.

DESCRIBE MTMM NEW STUFF HERE? Basically, words found through that project were imported and added to each foundation, with redundancies removed at the end.

The source article words were compiled into a dataset where they were matched up with their counterparts in the MFD along with their valence and a percentage of their occurrence. Therefore, for each article, the percentage of the number of \emph{harm/care} words occurring in the articles were calculated, and this process was repeated for each of the foundations. This procedure created five percentages that were included as the dependent variable for the following analyses. It should be noted the valence scores for the MFD words were \emph{z}-scored, so negative valence scores are possible.

@@

To analyze if news sources adhered to differences in word use based on their target audience, we utilized a multilevel model (MLM) to analyze the data. MLM is a regression technique that allows one to control for the repeated measurement and nested structured of the data, which creates correlated error (Gelman, 2006). Using the \emph{nlme} library in \emph{R} (Pinheiro, Bates, Debroy, Sarkar, \& Team, 2017), each foundation's weighted percentage was predicated by the political lean of the news source, using the individual news sources as a random intercept to control for the structure of the data.

The multilevel model did not indicate the presence of any significant or practical effect of political lean for any of the five moral foundations. The strongest effect size was observed for the \emph{authority/respect} foundation, but the effect was in the opposite direction from what was originally hypothesized - liberal sources tended to use more \emph{authority/respect} words than did conservative sources. Descriptive and test statistics, \emph{p}-values and effect sizes (Cohen's \emph{d}) can be found in Table 1.

>>>>

  \title{Moral Foundations of U.S. Political News Organizations}
    \author{William E. Padfield\textsuperscript{1}~\& Erin M. Buchanan, Ph.D.\textsuperscript{2}}
    \date{}

\shorttitle{MORAL NEWS}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Missouri State University\\\textsuperscript{2} }
\keywords{politics, morality, psycholinguistics}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

Morality and ethics have been of interest to thinkers, academics, and philosophers since antiquity. Starting chiefly in the twentieth century, a scientific approach to humans' understanding of morality emerged under the domain of psychology. Theories attempting to explain the development and application of people's moral intuitions built the foundation for the subfield of moral psychology. As the field developed, however, considerable debate has taken place regarding operational definitions of \enquote{morality.} Concerns regarding operationalization remain an issue in the field in the twenty-first century as researchers attempt to infer moral and political leanings from text and speech.

@@

As a discipline, modern moral psychology started in the late 1960s with Lawrence Kohlberg (Haidt \& Graham, 2007). Kohlberg's research popularized his theory of the development of moral reasoning. This theory establishes the steps of moral reasoning through which humans proceed as their cognitive structures assume higher levels of sophistication and nuance ({\textbf{???}}). Kohlberg borrowed from Jean Piaget's stages of cognitive development in which children progress from the sensorimotor through to the formal operations stage. Similarly, Kohlberg found people typically start with a \enquote{pre-conventional} understanding of morality during infancy in which children understand \enquote{right} and \enquote{wrong} purely in terms of how they interact with resultant experiences of rewards and punishment. Typically, people progress through several steps until they reach a \enquote{post-conventional} ethics. People who have reached the post-conventional stage are said to be able to weigh competing abstractions and reason their way to a conclusion that promotes justice based upon their \enquote{self-chosen ethical principles} ({\textbf{???}}). From Kohlberg's perspective, issues of justice and fairness comprise the foundation of morality (Haidt \& Graham, 2007). This view persisted until it encountered criticism in the early 1980s.

Kohlberg's conception of morality faced major scrutiny from psychologist Carol Gilligan. In 1982, Gilligan criticized Kohlberg's theory on the grounds that it focused solely on the moral concerns of men, and that it ignored those of women (Haidt \& Graham, 2007). Gilligan drew attention to purported differences in the ways men and women are taught to relate to self and others. She offered a historic argument contending women have traditionally filled roles related to caring and nurturing. She pushes back against Kohlberg's assumption that moral development replaces \enquote{rule of brute force,} as enforced by men, with the justice-based \enquote{rule of law.} According to Gilligan, this assumption implies women are less morally developed, owing to their absence both in masculine displays of violence as well as in enforcement of the law ({\textbf{???}}). Gilligan argues for the existence of a distinct, but equal development process that women and girls must undergo in order to develop their moral selves. Stark differences in the ways women are traditionally taught to interact with their social world cause them to develop ethical systems based upon their non-aggressive relationships with others. Gilligan thus asserted morality was built upon an alternative moral foundation: caring ({\textbf{???}}). This debate between competing conceptions of morality did not resolve until Gilligan and Kohlberg conceded the existence of two moral foundations: justice and caring (Haidt \& Graham, 2007). While this new direction in moral psychology appeared to represent a more inclusive outlook on the construct, these novel ideas would soon be challenged on the grounds of its apparent western-centric outlook.

@@

In their brief overview of the history of moral psychology, Graham, Haidt, and Nosek (2009) explained Shweder, Much, Mahapatra, and Park's objections to moral psychology as it stood in the late 1980s. Their criticism centered on the fact moral psychology concerned itself with issues regarding justice and individuals' rights. Such a system, they argued, did not account for moral concerns outside of the western world (Graham et al., 2009). Individually focused concerns can be grouped under an overarching \enquote{ethic of autonomy,} which was thought to be one of three ethics upon which humans base moral decisions. The other two ethics were the \enquote{ethic of community} (comprising one's duty to their family, tribe, etc.), and the \enquote{ethic of divinity}" - representing one's duty not to defile their God-given body and soul (Graham et al., 2009). In the 2000s, Haidt and Graham (2007) took this line of reasoning further in their assertion that moral psychology favored certain political ideologies over others.

Haidt and Graham settled on these specific foundations after the completion of a literature survey of research in anthropology and evolutionary psychology (Graham et al., 2011). The researchers attempted to locate virtues and morals corresponding to \enquote{evolutionary thinking.} For instance, the researchers cited Mauss' work on reciprocal gift-giving, which informed the establishment of the \emph{fairness/reciprocity} foundation. Additionally, evolutionary literature on disgust and its correlation to human behavior regarding food and sex informed the \emph{purity/sanctity} foundation (Graham et al., 2011). The researchers identified the five \enquote{top candidates} for the foundations of human cultures' morality (Graham et al., 2011).

@@

In order to capture language's role in moral and political reasoning, Graham et al. (2009) formulated the Moral Foundations Dictionary (MFD) in order to capture moral reasoning and justification as used in speech and text. The MFD is composed of 259 words, with around 50 words assigned to each of the five foundations. The researchers created a preliminary list of words that they believed would be associated with the five foundations. Then, using the Linguistic Inquiry and Word Count (LIWC; Pennebaker, Booth, \& Frances, 2007) computer program, they analyzed transcripts of liberal and conservative Christian sermons in order to obtain frequencies of the occurrence of words from the researchers' initial list. The researchers manually checked the results from LIWC in order to make sure the results make sense given the contexts and rhetorical devices used in the sermons, as word frequency analysis ignores sentence context. The researchers offered the following example from a Unitarian sermon as a demonstration of ambiguous statements requiring human verification: \enquote{Don't let some self-interested ecclesiastical or government authority tell you what to believe, but read the Bible with your own eyes and open your heart directly to Jesus} (Graham et al., 2009). This sentence added to the \emph{authority/respect} total in LIWC's analysis, but it appears to suggest that one should reject authority in this context. The researchers eliminated this sentence from the \emph{authority/respect} raw count on account of this discrepancy between the use of authority-related words and the speaker's clear intentions (Graham et al., 2009).

@@

When considering the issue of positive versus negative valence in MFD words, the question of how texts are analyzed vis-a-vis the MFD remains. How can raw percentage of MFD word occurrence capture the valence and focus of the writer or speaker? If 2\% of a politician's speech features positive words (i.e., \enquote{benefit} and \enquote{defend}) from the MFD \emph{harm/care} list, how can researchers be sure the level and nature of the speaker's \enquote{endorsement} of the foundation equals that of another politician whose speech contained negatively connoted MFD words from the \emph{harm/care} list? They would have equal endorsements as far as the numbers are concerned, but the words used and focus given are on opposite sides of the \emph{harm/care} spectrum.

@@

Participants saw the stimulus word and responded on each scale. The valence scale featured a smiling figure at one end (representing pleasantness) and a frowning figure at the other end (for unpleasantness). The arousal scale had a \enquote{wide-eyed} figure at one end with a sleepy figure at the other, representing stimulating and unstimulating respectively. Finally, the dominance scale featured a large figure, indicating the highest degree of control, at one end and a small figure, indicating a lack of control, at the other end (Bradley \& Lang, 1999). The end result of this procedure yielded affective norms along the three dimensions for 1,040 English words (Bradley \& Lang, 1999). ANEW represented an important first step in establishing affective norms for large numbers of English words. However, later researchers found the 1,040-word list to be limiting for a language consisting of thousands of words.

Warriner, Kuperman, and Brysbaert (2013) exponentially lengthened the list of words with affective norms to 13,915 English lemmas, the base forms of words without inflection (i.e., \enquote{watch} rather than \enquote{watched} and \enquote{watching}). The researchers recognized the importance of affective norms in several areas of study, including emotion, language processing, and memory (Warriner et al., 2013). They argue the list of words included in ANEW is sufficient for small-scale factorial research designs, but the list is \enquote{prohibitively small} for larger-scale \enquote{megastudies} that are common in psycholinguistic research today (Warriner et al., 2013).

In order to source a large number of lemmas for affective ratings, the researchers drew from several validated sources. These include the 30,000 lemmas with age-of-acquisition (average age at which a particular word is learned) ratings gathered by Kuperman, Stadthagen-Gonzalez, and Brysbaert (2012) as well as the content lemmas from the SUBTLEX-US corpus consisting of subtitles from various forms of visual media ({\textbf{???}}). This data collection resulted in the final list of 13,915 lemmas. Lists of 346-350 words were presented to participants recruited through the Amazon Mechanical Turk subject pool. Participants rated the words along one of the three dimensions, unlike the ANEW project in which participants rated each word along all three dimensions at once. The researchers used a nine-point scale similar to the one used by Bradley and Lang (1999) when collecting ratings for ANEW (Warriner et al., 2013).

The researchers noted several points of interest upon observing ratings. First, they found that valence and dominance ratings had a negative skew, indicating more words elicited feelings of happiness and control than their respective opposites. Also, when examining the relationship between valence and arousal ratings, the researchers found a U-shaped relationship. This U-shape indicates words with high degrees of positivity and negativity elicited higher arousal (Warriner et al., 2013). These observations along with the now-greatly expanded list of affective norms has been applied to several lines of inquiry in psycholinguistics.

Warriner and Kuperman (2015) utilized the new affective norms list in order to investigate the validity of the Pollyanna hypothesis, or the prevalence of a generally optimistic outlook in humans as reflected in language. The researchers were able to conclude the existence of a greater number of positive-valence English words in the list of 13,915 lemmas. Additionally, after observing token frequency in a number of text corpora, including SUBTLEX-US, COCA, BNC, TASA, and HAL, the researchers found that words with positive valence were also used more frequently (Warriner \& Kuperman, 2015). While the researchers concede the possibility of an acquiescence bias in ratings as a possible explanation for the observed positivity bias, this investigation represents one application of the Warriner et al. (2013) list in emotional studies.

In addition to applications in emotion research, the Warriner et al. (2013) norms have been utilized in cognitive research as well. One cognition-based study investigates the relationship between emotion and response latencies in word recognition. Kuperman, Estes, Brysbaert, and Warriner (2014) sought to use these new norms to fill in the knowledge gaps regarding variance in word recognition. The researchers drew several conclusions regarding emotion and word recognition (specifically in naming and lexical decision tasks). First, Kuperman et al. (2014) found slower decision-making and reading times in negative-valence words, faster times in neutral words, and even faster times in words with positive valence. The researchers also concluded that words causing higher arousal tend to have slower decision times than less-arousing words. They found valence had a stronger effect on recognition than arousal (both effects were independent, not interactive). They found an interaction between emotion and word frequency such that valence and arousal are more effective on lower frequency words than high frequency words. Finally, Kuperman et al. (2014) found a greater effect of valence and arousal on response latency for lexical decision tasks than for naming tasks (Kuperman et al., 2014). This research serves as further evidence that the Warriner et al. (2013) list can be used for research inquiries both within and without the field of psycholinguistics.

In the present studies, the researchers used the Warriner et al. (2013) list in order to denote the valence of the words appearing in the news articles scraped from the internet. Valence was considered as another independent variable and its relationship with the words comprising the Moral Foundations Dictionary were of chief interest to the researchers. The valence was used as a means to determine whether individual words in the MFD represented more positive aspects of their respective foundation or if they denoted a more negative aspect of the foundation. Specifically, valences were used to weight the MFD words by their relative degree of positivity or negativity. Incorporating word valence into a study involving the MFD is meant to alleviate some of the issues regarding the aforementioned ambiguity regarding the words in the Moral Foundations Dictionary.

@@

The use of language both to express and further an ideological goal has been documented in the techniques employed by candidates for political office in the U.S., Druckman, Jacobs, and Ostermeier (2004) considered political \enquote{issues} as communication that attempts to persuade constituents to vote for the candidates based on their strengths in matters of public policy. According to the researchers, \enquote{image} priming describes techniques deployed in order to sway votes based on favorable aspects of the candidate's behavior and personality (Druckman et al., 2004). The researchers investigated political issue and image priming on the part of candidates as implied by the disproportionate attention candidates paid to particular issues over others. The researchers found numerous examples of issue and image priming during the 1972 re-election campaign of Richard Nixon.

They linked the Nixon administration's awareness of the issues for which the president had public support to the issues he should emphasize (and prime) during the campaign. Likewise the researchers found evidence that Nixon's team was aware of negative evaluations of his warmth and trustworthiness, and thus took steps to prime his purportedly positive qualities, including strength and competence (Druckman et al., 2004). The researchers also cited research from Iyengar and Kinder (1987) suggesting the news media affected perceptions of President Jimmy Carter's competence by emphasizing (e.g.~priming) issues related to energy, defense, and the economy. This focus implies news media may contribute to Americans' perception of politicians based on where the media places emphasis.

There is a potential caveat regarding the validity of Druckman et al. (2004)'s findings: reproductions of several studies purporting to demonstrate social priming effects have failed to replicate the original results. ({\textbf{???}}) point out the distinction between perceptual and social (or goal) priming both in their operational definitions as well as their replicability. Perceptual priming often works through the inducement of a certain response from a related prime, as in, for example, semantic priming. Social (or goal) priming encompasses phenomena by which people exhibit complex behavioral changes subsequent to exposure to a prime. ({\textbf{???}}) point out well-known studies investigating social priming, including the use of elderly-related primes to induce slower walking speeds in participants. Studies investigating perceptual priming have been \enquote{directly replicated in hundreds of labs} ({\textbf{???}}). This does not appear to be the case for social priming, as argued by ({\textbf{???}}).

({\textbf{???}}) noticed the unusually large effect size values (Cohen's \emph{d}) reported by researchers studying social priming effects. The researchers reproduced two studies from ({\textbf{???}}) The first study attempted to prime participants by having them plot points on a Cartesian grid. The independent variable was priming condition and contained three levels: short, middle, and long distance. Those instructed to plot points further apart were hypothesized to express a higher degree of psychological distance regarding their family. The second study used the same priming conditions, but hypothesized that greater distance between points would prime participants to estimate fewer calories in unhealthy foods than those who were primed with shorter distances between points. ({\textbf{???}}) concluded those two studies from ({\textbf{???}}) held little validity while also casting doubt on the prevalence of social priming effects themselves, based on the inability of other researchers to replicate previously reported effects in this area.

While these concerns regarding the replication of social priming studies are valid and deserve further investigation, Druckman et al. (2004) does not purport to demonstrate a widespread effect of social priming on the American electorate. In other words, this reseach makes no claim to empirically supported priming effects. Rather, Druckman et al. (2004) chronicle the efforts on the part of the Nixon Administration to prop up the president's supposed strengths while downplaying his weaknesses. These tactics were deployed through the careful use of language in order to achieve the administration's political goals. As such, Druckman et al. (2004)'s research on Nixon serves as an example of language's potential utility in the propogation of desirable political opinions. The researcher's investigation of news media's focus on specific issues during the Carter Administration likewise provide an example of language as a potential conduit for the transfer of politically biased information. The idea that even 1970s news media could contain political biases is of particular interest to the current study, which investigates similar phenomena in contemporary news media.

Other research into news media suggests certain media outlets, at least indirectly, may have an effect on the voting records of representatives in Congress (Clinton \& Enamorado, 2014). Specifically, the researchers identified a pattern of declining support for President Bill Clinton's policies chiefly among Republicans in the House of Representatives after the Fox News Channel began broadcasting on cable and satellite systems in their respective districts. As Fox News was, at the time of its launch in 1996, the only outwardly ideological national news network, the researchers were able to track its spread across the country and observe voting records of members of Congress both before and after Fox News' arrival. The researchers concluded that members of Congress, excluding those newly elected at the time of Fox News Channel's emergence, attempted to anticipate resultant conservative-leaning shifts among their constituents by bolstering their conservative voting record before the next election (Clinton \& Enamorado, 2014).

@@

For Experiment 1, the researchers approached the study with the intention to answer a method question. That is, this portion of the current research was conducted in order to solidify the best method by which to analyze political news text under the Moral Foundations Theory framework while also alleviating some of the aforementioned valence problem observed in the Moral Foundations Dictionary. The researchers hypothesized the news sources genrally perceived as liberal leaning (\emph{NPR} and \emph{The New York Times}) would contain MFD words and valences indicating endorsements of the individualizing moral foundations (\emph{harm/care} and \emph{fairness/reciprocity}). Additionally, the researchers hypothesized the two sources generally perceived to be conservative leaning (\emph{Fox News} and \emph{Breitbart}) would feature MFD words and valences indicating equal endorsement of all five foundations.

@@

Political articles were collected from the websites of four notable U.S. news sources, a process known as web scraping. The sources were \emph{The New York Times}, \emph{National Public Radio (NPR)}, \emph{Fox News}, and \emph{Breitbart}. They were selected for their widespread recognition and the fact they are easily categorized (by the general public) according to perceived political lean. In general, \emph{The New York Times} and \emph{NPR} are perceived by many to have a liberal bias or lean. In contrast, Fox News and Breitbart are believed to have a conservative bias or lean (Mitchell, Matsa, Gottfried, \& Kiley, 2014). Political articles in particular were identified and subsequently scraped by including the specific URL directing to each source's political content in the \emph{R} script. For example, rather than scrape from nytimes.com, which would return undesired results (non-political features, reviews, etc.), we instead included nytimes.com/section/politics so that more or less exclusively political content was obtained. All code for this manuscript can be found at \url{https://osf.io/5kpj7/}, and the scripts are provided inline with this manuscript written with the \emph{papaja} library (Aust \& Barth, 2017).

@@

The article text was processed using the \emph{tm} and \emph{ngram} packages in \emph{R} in order to render the text in lowercase, remove punctuation, and fix spacing issues (Feinerer \& Hornik, 2017). The individual words were then reduced to their stems (i.e., \emph{abused} was stemmed to \emph{abus}). The same procedure was applied to the MFD words and the words in the Warriner et al. (2013) dataset. Using the Warriner et al. (2013) dictionary, the words making up each of the five foundations in the MFD were matched to their respective valence value.

Concurrent research by Jordan, Buchanan \& Padfield is assessing the validity of both the Moral Foundations Questionnaire and the Moral Foundations Dictionary through a multi-trait multi-method analysis of the two instruments using one sample. The instruments will be analyzed against one another, in order to test reliability, as well as against the Congressional Record in order to test predictive validity for political orientation. The researchers were able to identify a number of potential new words that, if added to the MFD, could comprise a dictionary with greater validity. Additionally, the researchers are investigating whether or not a reduced set of MFD words may likewise bolster the validity of the MFD. Depending on the results of that research, either an augmented or reduced Moral Foundations Dictionary may be utilized in this analysis rather than the original version.

The source article words were compiled into a dataset where they were matched up with their counterparts in the MFD along with their valence and a percentage of their occurrence. Therefore, for each article, the percentage of the number of \emph{harm/care} words occurring in the articles were calculated, and this process was repeated for each of the foundations. Words' percent occurence were multiplied by their \emph{z}-scored valence. Valences were \emph{z}-scored in order to eliminate any ambiguity regarding the direction of the valence. Positive values indicate positive valence, and negative values indicate negative valence. Words were categorized in accordance to their MFD affiliation, creating a weighted sum for each moral foundation.

@@

To analyze if news sources adhered to differences in word use based on their target audience, we utilized a multilevel model (MLM) to analyze the data. MLM is a regression technique that allows one to control for the repeated measurement and nested structured of the data, which creates correlated error (Gelman, 2006). Using the \emph{nlme} library in \emph{R} (Pinheiro, Bates, Debroy, Sarkar, \& Team, 2017), each foundation's weighted percentage was predicated here by the political lean of the news source, using the individual news sources as a random intercept to control for the structure of the data.

The multilevel model did not indicate the presence of any significant or practical effect of political lean for any of the five moral foundations. The strongest effect size was observed for the \emph{authority/respect} foundation, but the effect was in the opposite direction from what was originally hypothesized - liberal sources tended to use more \emph{authority/respect} words than did conservative sources. Descriptive and test statistics, \emph{p}-values and effect sizes (Cohen's \emph{d}) can be found in Table 1. Based on the weighted percent values for the five foundations, the researchers observed that MFD words seem to make up a small portion of the article text. Furthermore, the observed percentages and means appear to indicate a generally positive endorsement of all five foundations across both liberal and conservative sources.
